---
title: "Exercises_0202"
author: "Cory Suzuki"
date: "2/2/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Exercise 3
Explain whether each scenario is a supervised or unsupervised learning task.

1. A cancer researcher want to identify subgroups among 100
cancer samples in order to obtain a better understanding of the
disease. The number of subgroups is unknown.

This problem is an unsupervised learning problem because the number of k subgroups is unknown. This would use the clustering problem algorithm.

2. A cancer researcher want to build a model to classify 100
samples into one of the two groups: benign or malignant, using
the features: age, genetic factors, family history of cancer.

This problem is a supervised learning problem because we are classifying age, genetic factors, and family history which are quantitative responses. This would fall under the regression algorithm.

##Exercise 4
Explain whether each scenario is a classification or regression problem and indicate whether we are most interested in inference or prediction. Finally, provide the sample size n and the number of features p.

1. We collect a set of data on the top 500 firms in the US. For
each firm we record profit, number of employees, industry and
the CEO salary. We are interested in understanding which
factors affect CEO salary.

This is a regression problem since we are asked to understand the quantitative relationship between the variables of record profit, number of employees, industry salary and how they affect the response of CEO salary. Thus we are most interested in inference, with n=500 and p=3 features.

2. We are considering launching a new product and wish to know
whether it will be a success or a failure. We collect data on 20
similar products that were previously launched. For each
product we have recorded whether it was a success or failure,
price charged for the product, marketing budget, competition
price, and ten other variables.

This is a classification problem because we are asked to qualitatively analyze the effectiveness of the product either being a success or a failure. Here we are interested in making a prediction regarding the product's functionality. Here, n=20 and p=13 (three known variables and ten other variables).

3. We are interested in predicting the ed % change in the
USD/Euro exchange rate in relation to the weekly changes in
the world stock markets. Hence we collect weekly data for all
of 2012. For each week we record the % change in the
USD/Euro, the % change in the US market, the % change in
the British market, and the % change in the German market.

This is a regression problem as we are analyzing how the variables of the US market, the German market, and the British market affect the % change in the US/Euro exchange market which is quantitative. We are most interested in making a prediction between known variables and an unknown response. Here, n=52 weeks in 2012 and p=3.

##Exercise 1
1.This code chunk loads the GGally, MASS, and tidyverse packages and imports the dataset for the Boston criminal statistics. The data is displayed as a tibble.
```{r}
library(GGally)
library(MASS)
library(tidyverse)
data(Boston)
as_tibble(Boston)
help(Boston)
```

2.This code chunk builds a simple linear regression model of the Boston dataset where y=medv and x=lstat.
```{r}
lin_mod = lm(medv ~ lstat, data = Boston)
summary(lin_mod)
```
The coefficient estimate for Beta hat of lstat is -0.950. We interpret this as: For every one percent increase of the lower status of the population, on average, it is associated with a 0.950 decrease in the median value of owner-occupied homes in \$1000s.

3.This code chunk graphs a scatterplot of the linear regression model performed on the Boston dataset from part 2, with y=medv and x=lstat.
```{r}
ggplot(data = Boston, mapping = aes(x = lstat, y = medv)) +
geom_point() +
geom_smooth(method="lm", formula = y ~ x) +
ggtitle("Median Value of Owner Occupied Homes  
         vs Lower Status of the Population (linear model)") +
theme_bw()
```