---
title: "Lecture_2_Exercises"
author: "Cory Suzuki"
date: "1/31/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exercise 1 

##1.This code chunk loads the MASS Boston library, dplyr, and GGally. It uses help() to identify what each variable is in the Boston package.
```{r}
library(MASS)
library(tidyverse)
library(dplyr)
library(GGally)
data(Boston)
help(Boston)
```

##2.This code chunk uses select() to select the following subset of variables from the Boston dataset.
```{r}
library("tidyverse")
select = dplyr::select
Boston_subset = select(Boston, crim, indus, nox, age, lstat) %>%
as_tibble()
print(Boston_subset)
```
There are 506 rows, 5 columns, and 5 features in this subset.

##3. This code chunk uses ggpairs() to make a scatterplot matrix of the data subset from part 2.
```{r}
library("GGally")
ggpairs(Boston_subset)
```

##4. This portion is a written response to the proposed question: According to the scatterplot matrix in part 3, which predictor has the strongest correlation with per capita crime rate per town?

According to the scatterplot matrix in part 3 of the Boston data subset, the predictor of lower status of the population in percent has the strongest correlation with per capita crime rate per town (r=.456).

Exercise 2
What is the goal of each of the following cases: prediction or inference? Describe each briefly.

1. We want to know whether a patient, hospitalized due to a heart
attack, will have a second heart attack, on the basis of the
demographic, diet and clinical measurements for that patient.

This case's goal is a prediction whether a person will have a second heart attack(response) based on the demographic, diet, and clinical measurements of the individual(predictor).

2. We want to identify and quantify the influence of each
important predictor for heart attack, based on demographic,
diet and clinical variables.

This case's goal is an inference since it aims to identify/select which predictors are responsible for the response.

3. We want to know the price of a stock in 6 months from now,
on the basis of company performance measures and economic
data.

This case's goal is a prediction since it wants to find out the price of the stock (the output) based on performance measures and economic data (the inputs).
