---
title: "Exercises_0214_0216"
author: "Cory Suzuki"
date: "2/14/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Exercise 1

1. This code performs a univariate logistic regression on the Default dataset with the response variable as default and income as the predictor.
```{r}
library(tidyverse)
library(ISLR)
data(Default, package="ISLR")
m1 = glm(default ~ income, family = "binomial", data = Default)
summary(m1)
```
The logistic equation in terms of log odds is:
$log(p/1-p) = -3.094e+00-8.353e-06 \hat{\beta}_1$.

2. The expected change in odds of default is by `exp(-8.353e-06)` units.

3. This code chunk calculates $P(X | x=33)$.
```{r}
(exp(-3.094e+00-8.353e-06*33))/(1+(exp(-3.094e+00-8.353e-06*33)))
```

4. This code chunk calculates $P(X| x=33)$ using the predict glm function.
```{r}
help(predict.glm)
newd = data.frame(income = 33)
predict(m1, newd, type = "response")
```

##Exercise 2
1.
```{r}
library(MASS)
library(tidyverse)
data(Boston)
df2 = Boston %>%
  mutate(crim_binary = ifelse(crim >= 0.26, 1, 0))
df2
```

2.
```{r}
log_mod_bos = glm(crim_binary ~ indus + nox + age + lstat, family = "binomial", data = df2)
summary(log_mod_bos)
```
3. Based on the summary function, the log odds is:
$log(p/(1-p)) = -18.044264-0.076756(indus)+34.528941(nox)+0.005199(age)-0.010058(lstat)$

4. We interpret the beta hat coefficient for lstat as: For every proportion of non-retail business acres per town, every parts per 10 million of nitrogen oxides concentration, and every proportion of owner-occupied units built prior to 1940, the one percent increase in the lower status population is associated with a 6.6% decrease of the log odds of the criminal binary response variable.

5. This code chunk computes the 95% confidence interval for the multiple logistic regression model we have created.
```{r}
help(Boston)
confint(log_mod_bos, level = 0.95)
```
We interpret the confidence interval of the beta hat coefficient for lstat as: We are 95% confident that the one percent increase in the lower status population is associated with a decrease of 6.6% or a 4.4% increase of the log odds of the criminal binary response variable.