---
title: "Exercises_0207"
author: "Cory Suzuki"
date: "2/7/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Exercise 2
1. This chunk of code loads the MASS, GGally, and tidyverse libraries and imports the Boston data set.
```{r}
library(MASS)
library(GGally)
library(tidyverse)
data(Boston)
help(Boston)
```

This chunk of code builds a simple linear regression model of the Boston data set with y=medv and x=lstat. We also determine a 90% confidence interval of our model.
```{r}
lin_regn = lm(medv ~ lstat, data = Boston)
confint(lin_regn,level = 0.90)
```
The 90% confidence interval for beta nought (the slope estimate) is [-1.014, -0.886]. We interpret this as a conclusion that for each one percent increase of the lower status population, there will be an average decrease in the median value of owner-occupied homes between \$1,014 and \$886.

2. This code chunk prints a summary of the simple linear regression model and here we analyze the hypothesis test of our beta nought and get our R-squared value.
```{r}
summary(lin_regn)
```
According to the printed summary, we reject the null hypothesis and conclude that for the Boston data set that there is a significant relationship between the median value of owner-occupied homes in \$1000s and the lower status of the population since our p-value is less than 0.05. The R-squared of the linear model is 0.543.

##Exercise 3
1. This code chunk creates another object instance of the Boston linear regression model with the addition of two more predictors, namely age and pupil-teacher ratio.
```{r}
lin_regnv2 = lm(medv ~ lstat + age + ptratio, data = Boston)
summary(lin_regnv2)
```

2. This code chunk makes a new data frame and assigns values to the predictor variables lstat, age, and ptratio. It also calls the predict() function and sets the level to 0.95.
```{r}
newd = data.frame(lstat = 13, age = 69, ptratio = 18)
predict(lin_regnv2, newd, interval = "prediction", level = 0.95)
```
The predicted medv (median value of owner-occupied homes in \$1000s) value is 22.765.